Hysteresis in human computation:\\ how one task affects another

### 2014-12-03 docs/publications/2014.09_SCIENCE/man.tex
small edit

Faced with the scale and complexity of the data revolution, researchers and
industry are increasingly relying on \textit{human computation}.  Microtask
platforms, like Amazon Mechanical Turk, act like a human compute server,
making it possible to access human cognition quickly and inexpensively. 
This has lead to broad adoption across a spectrum of disciplines
in the humanities and computer science, as well as in research into medical 
imaging technologies and other applications where expert opinion would 
normally be sought.  Here we show that an inherent
feature of microtask platforms, the repetition of similar tasks,
acts as a strong, but as yet unexplored source of bias.  
These \textit{intertask effects}, rival the
severity of overtly framing a task's purpose.  
Using image-labeling as a canonical microtask, we found that prior 
tasks influences the vocabulary that a worker uses when labeling 
subsequent images in profound an nuanced ways.  Prior tasks not only 
alter what a worker focuses on, 
but also modulate the richness and specialization of her vocabulary.  
While intertask effects can be a source of systematic bias, our
results suggest that, when properly controlled, they can be leveraged
to hone worker focus and specificity.  Though we focused on microtask 
platforms due to their emerging prominence, our results apply to  
citizen science, crowdsourcing, and any application involving repetitive 
tasks relying on human judgment.
\end{sciabstract}

>>>>

Faced with the scale and complexity of the data revolution, researchers and
industry are increasingly relying on \textit{human computation}.  Microtask
platforms, like Amazon Mechanical Turk, act like a human compute server,
making it possible to access human cognition quickly and inexpensively. 
This has lead to broad adoption across a spectrum of disciplines
in the humanities and computer science, as well as in research into medical 
imaging technologies and other applications where expert opinion would 
normally be sought.  Here we show that an inherent
feature of microtask platforms, the repetition of similar tasks,
acts as a strong, but as yet unexplored source of bias.  
These \textit{intertask effects} rival the
severity of overtly framing a task's purpose.  
Using image-labeling as a canonical microtask, we found that prior 
tasks influence the vocabulary that a worker uses when labeling 
subsequent images in both profound an nuanced ways.  Prior tasks not only 
alter what a worker focuses on, 
but also modulate the richness and specialization of her vocabulary.  
While intertask effects can be a source of systematic bias, our
results suggest that, when properly controlled, they can be leveraged
to hone worker focus and specificity.  Though we focused on microtask 
platforms due to their emerging prominence, our results apply to  
citizen science, crowdsourcing, and any application involving repetitive 
tasks relying on human judgment.
\end{sciabstract}
